Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-05-11 paper/manuscript.Rmd
q3
conclusion

We surmise that as competition among conspecifics increased along with
population size, density-dependent habitat selection resulted in expansion of
the population to nearby islands. Following Ideal Free Distribution theory,
caribou should swim to new islands when the average fitness of individuals on
the starting island exceeds the density-fitness equilibrium [@Morris1987]. While
this is an ultimate explanation, our data suggest caribou only remain on smaller
islands for a few days at a time, so the fitness equilibrium remains a
theoretical construct. More likely, however, is that fine-scale competition,
density-dependent habitat selection, and forage depletion drives individuals to
periodically swim between islands. Given the presumed low density of coyotes on
Fogo Island and the minimal risk they pose, predation risk does not appear to be
a major driver of swimming. Our observations add to the evidence that caribou
can, and occasionally do, swim in the ocean [e.g. @Miller1995; @Miller2002;
@Jeffery2007; @Ricca2012]. We suggest that swimming is likely more common than
previously thought for caribou living on oceanic islands, and that forage
limitation and the associated density-dependent habitat selection is an ultimate
explanation for this phenomenon. Although our inference is limited to
observations, islands appear to represent discrete foraging patches for
terrestrial animals that can influence fitness via increased foraging
opportunities. We also present several hypotheses related to the nuances and
further variation within the idea that density-dependent habitat selection
governs swimming behaviour, and posit that above and beyond these additional
factors, individuals can vary in their evaluation of costs and benefits of
movement between these habitat patches.

>>>>

We surmise that for Fogo Island, as competition among conspecifics increased along with
population size, density-dependent habitat selection resulted in expansion of
the population to nearby islands. Following Ideal Free Distribution theory,
caribou should swim to new islands when the average fitness of individuals on
the starting island exceeds the density-fitness equilibrium [@Morris1987]. While
this is an ultimate explanation, our data suggest caribou only remain on smaller
islands for a few days at a time, so the fitness equilibrium remains a
theoretical construct. More likely, however, is that fine-scale competition,
density-dependent habitat selection, and forage depletion drives individuals to
periodically swim between islands. Our focal observations add to the evidence that caribou
can, and occasionally do, swim in the ocean [e.g. @Miller1995; @Miller2002;
@Jeffery2007; @Ricca2012]. We suggest that swimming is likely more common than
previously thought for caribou living on oceanic islands, and that forage
limitation and the associated density-dependent habitat selection is an ultimate
explanation for this phenomenon. Although our inference is limited to
observations, islands appear to represent discrete foraging patches for
terrestrial animals that can influence fitness via increased foraging
opportunities. We also present several hypotheses related to the nuances and
further variation within the idea that density-dependent habitat selection
governs swimming behaviour, and posit that above and beyond these additional
factors, individuals can vary in their evaluation of costs and benefits of
movement between these habitat patches.
